A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines

### 2019-07-28 latex/paper.tex

@@

\begin{abstract}
% Topic
As the amount of data increases and is easier to access, Big Data processing becomes
critical in neuroimaging.
% Problem
This is an issue as the currently used frameworks are specialized for neuroimaging
and not Big Data concerns. This can lead to performance decrease or limit the
research we can perform.
% Relevance
The usage of Big Data framework is beneficial to this problem. The state-of-the-art
general purpose Big Data framework Spark has its code base written in Scala, while
our laboratory mostly uses Python. This can be problematic since it is harder to port
our pipelines to the framework. Also, theoretically, it leads to performance decrease.

% Approach
We propose to use Dask since it provides native Python parallelism to pipelines while
providing support to familiar API from the Python's scientific ecosystem.
Unfortunately, there are few comparisons between Spark and Dask; especially in
neuroimaging. Moreover, these studies were done when Dask was still an immature
framework which makes those studies unfair nowadays. This is our motivation to
compare the latest version of Dask with Spark.
% Methods (one sentence)
To evaluate the frameworks, we focus on their performance and their scheduler.

% Key Impact
Our study demonstrates the potential of using Dask as the framework to build
neuroimaging pipelines.
\end{abstract}

@@

%Similarities
Spark and Dask both offer in-memory computing, data locality, and lazy evaluation;
which is common for Big Data frameworks. Both their schedulers operate dynamically,
which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
also provide rich, high-level programming APIs, and support a variety of
infrastructure schedulers, such as Mesos, YARN, or HPC clusters (Dask only). Over
these similarities, the frameworks have differences.

@@

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine.
While its core is written in Scala, Spark provides additional APIs in Java, R
and Python (PySpark). Its main abstraction, the Resilient Distributed Dataset
(RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements. It 
achieves fault tolerance through recording the data lineage, the sequence of 
operations required to produce the data. The RDD is the basis of Spark's other
data structures, namely, DataFrames
and Datasets. Datasets are similar to RDDs but additionally use Spark SQL's
optimized execution to further improve performance. DataFrames, used to process
tabular data, Datasets where the data is organized into named-columns. While 
the DataFrame API exists in all the available language APIs, Datasets are
limited to Scala and Java.

As data transfers in BigData workflows are an important source of performance
bottlenecks, Spark incorporated
data locality, and additionally, introduced in-memory computing to the Big Data 
frameworks. Data locality, introduced by MapReduce~\cite{mapreduce}, schedules
tasks as close as possible to where the data is storage. In-memory computing ensures
data is maintained in memory whenever possible, as writting large amounts data to
slower storage devices may be costly. To reduce any unecessary communication
and computation, Spark also included lazy evaluation, which builds the entire
task graph prior to execution to determine what needs to be computed.

Spark is compatible with three different schedulers: Spark
standalone, YARN~\cite{vavilapalli2013apache} and Mesos~\cite{hindman2011mesos}. 
The Spark standalone is a simple default scheduler built into Spark. YARN is a
more complex scheduler primarily designed to schedule Hadoop-based workflows. 
Mesos, on the other hand, is not limited to Hadoop and can be used to schedule
a variety of different workflows. As scientific researchers would likely be
executing their workflows in HPC environments with neither YARN or Mesos
installed, we limit our focus to Spark's Standalone scheduler.

The Spark Standalone scheduler is composed of three main processes: the
\emph{master}, the \emph{workers} (\emph{slaves}, in Spark) and the
\emph{driver}. The \emph{master} coordinates resources provisioned by
\emph{worker} processes on the cluster. The application is submitted to the
\emph{driver} that in turn requests workers to the master and dispatches tasks
to them. A job is divided into stages to be excuted in a different process onto
the workers. Each stage's operation is represented as a high-level task in the computation
graph. The Spark standalone scheduler uses a FIFO (First-In-First-Out) job
scheduling policy. The Spark standalone scheduler has two
execution modes: client mode, where the driver runs in a dedicated process outside
of the Spark cluster, and cluster mode, where the driver runs in a worker. We used
client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.

Python is commonly selected as the programming language of choice in scientific
communities, especially for our use case of neuroscience, where numerous specialized
Python libraries exist to study the data. While serialization of 
Python to Java may lead to significant overheads, we chose to focus on Spark 
Python API due to its suitability for neuroimaging research.

\subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
in the scientific Python ecosystem. Like Spark, Dask provides
in-memory computing, data locality, lazy evaluation, and fault tolerance by
means of data lineage. In contrast, Dask's core is purely in Python, eliminating
the cost of serialization to Java. Furthermore, it is lightweight and modular,
allowing users to only install the components they require.

Dask has five main data structure APIs:
Array, Bag, DataFrame, Delayed, and Futures. A Dask
\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
processing of large arrays. It provides a distributed clone of the popular NumPy
library. Similar to Spark's RDD, a Dask
\href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects. It offers a programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. A
\href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
as they support arbitrary tasks, but they operate in real-time rather than lazily.
All Dask data structure APIs, with the exception of Dask Arrays and Dask
DataFrames, were used in our experiments.

While Python multithreading is limited by the
\href{https://docs.python.org/3/glossary.html#term-gil}{Global Interpreter Lock (GIL)}
as a result of CPython's non thread-safe memory management, Dask attempts to
bypass the GIL whenever possible and use multithreading to provide parallelism.
With the exception of Dask Bag, all APIs, by default, use the local multithreaded
scheduler. Dask Bag instead relies on the local multiprocessing scheduler.

@@

 Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
 used Dask's default configuration that uses all the available memory on the
 instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
 Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave 5~GB
 for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
 default configuration for the master. We used the default confirguration for worker
 memory management: at 60\% it spill data to disk, and 50\% of that amount is
 reserved for storage and is immune to eviction.

 One cloud instance did not host any worker and had a \SI{2}{\tera\byte} disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, and for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. For both Spark and Dask, spilled data is evicted to the NFS.

@@

We also used the dataset provided by the Consortium for Reliability and
Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
available on \href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and
functional images of 1,397 subjects acquired in 29 sites. We used all 3491 anatomical
images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} by image on
average).

@@

%% Incrementation
\subsubsection{Incrementation}
We used an adaptation of the image incrementation pipeline used
in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
The application reads blocks of the BigBrain image from the shared file
system, increments the intensity value of each voxel by 1 to avoid caching
effects, sleeps for a configurable amount of time to emulate more
compute-intensive processing, repeats this process for a specified number of
iterations, and
finally writes the result as a NifTI image back to the shared file system.
This application allows us to study the behavior of the engines when all
inputs are processed independently, in a map-only scenario (see
Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
independent subjects in parallel.

@@

 Our second application calculates the histogram of the BigBrain image (see
 Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared file
 system, calculates intensity frequencies, aggregates the frequencies across the
 blocks, and finally writes the resulting histogram to the shared file system as a
 single \SI{766}{\kilo\byte} file. This application implements a typical map-reduce
 program, where the final result is obtained from all the input blocks (see
 Fig.~\ref{fig:tg-histo}). This application requires data shuffling, incurring 
 additional performance costs compared to Algorithm~\ref{alg:incrementation}.
 The total amount of shuffled data is, however, limited to
 \SI{2.62}{\mega\byte} per blocks as it only consists of image histograms.
Two implementation are studied: pure Python and NumPy.

@@

We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example} application
available on GitHub. This application can also be broken down into a map and
reduce phase. The map phase,
also called participant analysis, extracts the brain tissues of a subject's 3D MRI
using the popular FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting image
(\SI{2.47}{\mega\byte} on average) to the shared file system. The reduce phase, also
called group analysis, computes the volume of each brain and returns the average
volume, shuffling a total of \SI{8.6}{\giga\byte} image data.

While Incrementation and Histogram were implemented directly in Python, BIDS App
example requires an external library and is distributed as a Docker container image
(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
given its popularity for HPC and reproducibility, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
The image was preloaded on the nfs and singularity version 3.2.1\-1.el7 was installed
on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
export the path for fsl.

@@

We varied four parameters in our experiments, as shown in
Table~\ref{tab:param}. We varied (1) the number of workers to assess the
scalability of the engines, (2) the number of BigBrain blocks in
Incrementation and Histogram to measure the effect of different IO patterns
and parallelization degrees, and (3) the number of iterations and sleep
delay in Incrementation to evaluate the effect of job length.
It should be noted that increasing the number of blocks or iterations also
increases the total compute time of the application for a given sleep
delay. To avoid any potential external bias such as background load on the
network, we ran the experiments in a randomized order and cleared the page
cache of each worker before each execution.

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead across all application tasks. 
The overhead calculation for each CPU is the end time of its last processed task
minus the start time of the application; the summation of those results is the total
overhead.

@@

For all engines, the makespan is far from decreasing linearly with the
number of workers. Note that each worker has 8 CPUs. The makespan
even increases between 4 and 8 workers. This is due to the high impact of
data transfers and engine overhead on the application.

@@

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's in particular as the number of workers increase.
Within Dask, Delayed and Futures have a higher overhead than Bag. However,
overhead differences are compensated by an increase in data transfer time,
as a reduced overhead increases the concurrency between data transfers.

@@

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks, Delayed and Futres have much higher overhead. This is because they are
thread-based and use multiple threads to compute a block which affect the calculation
of the overhead. Once again, the data transfer time reduces with more blocks however
the overhead time increases by a similar amount. This is not observed for 30 blocks
as the workers are not used at full capacity, i.e., some threads are idle. Finally,
the variability of the overhead increases with the number of blocks, which explains
the makespan variability mentioned previously.

@@

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
for different sleep delays. Overall, all engines again perform the same and scale
well with task duration. Spark is initially slower than the Dask APIs, however, it is
faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two but it is not considerable.

@@

%%% HISTOGRAM %%%
\subsection{Histogram: Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
amounts of workers. Spark is tremendously faster than Dask APIs. The difference
narrows as the number of workers increases but stays significant. Between Dask APIs
there is no substantial difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
shown. The computation time is significantly larger for Dask than for Spark. This is
potentially due to Python's GIL preventing Dask to parallelize the computation on
multiple threads. Overall, the IO and overhead are comparable for all engines.

On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
when their number of workers varies, except for 8 workers where it increases slightly.
On the other hand, Spark total execution time keeps increasing; especially at 8
workers. This is because Dask engines benefit more from additional workers.

@@

\subsection{Histogram: Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block size. Again,
Spark is not compared at 30 blocks due to its \SI{2}{\giga\byte} task limitation.
Once more, Spark is significantly faster than Dask engines and Bag is slightly faster
than Delayed. Overall, the engines do not react to change in block size.

In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
Note that at 30 blocks the worker's resources are not fully used since the workers
can process up to 64 tasks in parallel. Moreover, the overhead for 30 blocks is
erroneous due to a lower amount of block than available threads. This leads to thread
based engines to have higher overhead. Given that,

Spark has a much lower compute time which makes it significantly faster than Dask
engines. Overall, for all engines, lower block size results in lower IO time but
increases the overhead time. Again, this is due to large overhead desynchronizing task
thus reducing the NFS bottleneck.

@@

\subsection{Histogram: Gantt chart}
Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline Histogram
experiment. Spark overhead is distributed through all the execution while Bag
overhead happens mostly when a thread read subsequent blocks and Delayed overhead is
concentrated after reading all the blocks.

The reduce is done differently for the engines. Spark and Delayed compute all the
reduce tasks on one process (thread), however, Spark only computes it on the driver
process. On the other hand, Bag computes the reduce tasks on different workers and
sends the information to the driver process afterward.

@@

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
Overall, the makespan for all engines does not change considerably with an increase
in the number of workers. This is because the engines do not scale well. There is no
substantial difference between all engines, although, Delayed is slightly faster.

@@

\subsection{Histogram (NumPy): Number of blocks}
Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
different block size. Again, Spark is not compared at 30 blocks due to its
\SI{2}{\giga\byte} task limitation. Overall, there is no substantial difference for
all engines and block sizes.

In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
Once more, since Delayed is thread-based its overhead at 30 blocks is seems much
higher that it really is. Considering the overhead error and that at 30 blocks the
workers only use half their core, the total time is substantially the same for all
block size. This is due to lower block sizes having a lower IO time, however, it
balances out by the resulting increase in overhead.

@@

\subsection{Histogram (NumPy): Gantt chart}
Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline for the
Histogram (NumPy) experiment. The overhead for Spark and Bag is mostly located
between a read and compute task or compute and read task. On the other hand, Delayed
overhead is dispersed between all types of tasks. This causes Delayed to overlap more
compute and read tasks simultaneously among the different workers.

@@

%%% BIDS EXAMPLE %%%
\subsection{BidsApp example: Number of workers}
Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when varying
the amount of workers. Overall, there is no substantial difference between the
engines. The makespan scales well with the increase in workers, however, the scaling
reduces considerably when reaching 8 workers. This is due to an increase in overhead
and of the IO requirement increasing with the number of workers --- more tasks are
read concurrently and the application becomes more data-intensive.

@@

\subsection{Effect of Network File System}
We used an NFS as it is a common approach to store a large amount of data. However,
it limits the available bandwidth since all workers interact with one disk. This
causes data-intensive applications to take a significant amount of time for IO. As we
expect, this effect is amplified when too many tasks simultaneously try to operate
with the NFS at the same time as
Fig.~\ref{fig:inc_gantt},\ref{fig:histo_gantt},\ref{fig:histo_np_gantt},\ref{fig:bids_gantt}
shows. We could use a parallel shared file system, like Lustre \MD{cite lustre if not
done already}, to reduce the impact of operations with the disk, however, we would
quickly be limited by the network's bandwidth.

>>>>

%\setlength{\textfloatsep}{7pt}

@@

\begin{abstract}
% Topic
Recent increases in data volume and improved accessibility have 
led neuroimaging into the Big Data era. 
% Problem
While existing neuroimaging pipeline frameworks typically do not attempt to
limit data transfers, Big Data frameworks remain seldom used.
% Relevance
In this paper, we compare two popular Big Data frameworks with Python APIs,
Apache Spark and Dask, for their runtime performance in processing neuroimaging
pipelines.
% Approach
To evaluate performance differences we used two synthetic, Incrementation
(map-only) and Histogram (map-reduce), pipelines and real neuroimaging map-reduce pipeline (BIDS App Example). For both our 
synthetic pipelines we varied the number of workers and block size using the 
76~GB BigBrain image. We additionally varied the number of iterations and task
duration for our Incrementation pipeline. 39~GB of anatomical images from the
CoRR were used for the BIDS App Example pipeline.
% Methods (one sentence)
All experiments were executed on 8 instances of Compute Canada's Arbutus cloud
with 8~vCPUs and 30 GB of RAM all connected to a 2~TB shared NFS. We evaluated
PySpark's RDD against Dask's Bag, Delayed and Futures.
% Key Impact
Our result show that despite the differences between the two frameworks, both
frameworks perform the comparably. However, Dask pipelines risk being limited by
Python's GIL depending on task type and cluster configuration. In all cases, 
the major limiting factor was data transfer. While either framework is suitable
for neuroimaging pipelines, more effort needs to be placed in reducing data 
transfer time.
\end{abstract}

@@

%Similarities
Spark and Dask both offer in-memory computing, data locality, and lazy evaluation;
which is common for Big Data frameworks. Both their schedulers operate dynamically,
which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
also provide rich, high-level programming APIs, and support a variety of
infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
these similarities, the frameworks have differences.

@@

\subsubsection{Apache Spark} Apache Spark is a widely-used 
general-purpose Big Data engine.
Its main abstraction, the Resilient Distributed Dataset
(RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements. It 
achieves fault tolerance through the recording of data lineage, the
sequence of operations used to modify the original data. The RDD is the basis
of Spark's other data structures, namely, Datasets
and DataFrames. Datasets are similar to RDDs but additionally use Spark SQL's
optimized execution to further improve performance. DataFrames, used to process
tabular data, are Datasets where the data is organized into named-columns.
While the DataFrame API exists in all the available language APIs, Datasets are
limited to Scala and Java.

As data transfers in Big Data workflows are an important source of performance
bottlenecks, Spark incorporates the concepts of data locality and in-memory 
computing. Data locality, introduced by MapReduce~\cite{dean2008mapreduce},
schedules
tasks as close as possible to where the data is storage. In-memory computing
ensures data is maintained in memory whenever possible, as writing large 
amounts data to slower storage devices may be costly. To reduce any unnecessary
communication and computation, Spark also included lazy evaluation, which
builds the entire task graph prior to execution to determine what needs to be
computed.

% Didn't mention Kubernetes as it's still experimental
Spark is compatible with three different schedulers: Spark
standalone, YARN and Mesos.
The Spark standalone scheduler is a simple default scheduler built into Spark.
In contrast, the YARN scheduler is primarily designed to schedule Hadoop-based 
workflows, whereas Mesos can be used to schedule
a variety of different workflows. As scientific researchers would likely be
executing their workflows in HPC environments with neither YARN or Mesos
installed, we limit our focus to Spark's Standalone scheduler.

The Spark Standalone scheduler is composed of three main processes: the
\emph{master}, the \emph{workers} (\emph{slaves}, in Spark) and the
\emph{driver}. The \emph{master} coordinates resources provisioned by
\emph{worker} processes on the cluster. The application is submitted to the
\emph{driver} that in turn requests workers to the master and dispatches tasks
to them. A job is divided into stages to be executed in a different process onto
the workers. Each stage's operation is represented as a high-level task in the
computation graph. The Spark standalone scheduler uses a FIFO 
(First-In-First-Out) job scheduling policy. The Spark standalone scheduler has
two execution modes: client mode, where the driver runs in a dedicated process
outside of the Spark cluster, and cluster mode, where the driver runs within
a worker process. Our experiments use client mode as cluster mode is not
available in PySpark.

Python is commonly selected as the programming language of choice in scientific
communities, and in particular, for our use case of neuroscience, where
numerous specialized Python libraries exist to study the data.
While serialization of Python to Java may lead to significant overheads, we
chose to focus on Spark Python API due to its suitability for neuroimaging
research.

We used Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Python-based Big Data engine that is becoming
increasingly popular in the scientific Python ecosystem. Like Spark, Dask avoids data transfers and needless computation and communication through
in-memory computing, data locality, lazy evaluation. Dask worklflows can 
further reduce data transfer costs by leveraging multithreading whenever 
communication is not bounded by Python's GIL. Dask relies on data 
lineage to achieve fault tolerance, however, unlike Spark, does not require
operations to be coarse grained. Furthermore, Dask is lightweight and modular,
allowing users to only install the components they require.

Dask has five main data structure APIs:
Array, Bag, DataFrame, Delayed, and Futures. A Dask
\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
processing of large arrays. It provides a distributed clone of the popular NumPy
library. Similar to Spark's RDD, a Dask
\href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
collection of Python objects. It offers a programming abstraction similar to the
\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. A
\href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
composition of
\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas Dataframes} used to process a large amount of tabular data.
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports
arbitrary tasks
that do not fit in the Array, DataFrame or Bag APIs. Finally,
\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
Delayed in that they support arbitrary tasks, but they operate in real-time
rather than lazily. With the exception of Dask Bag, all APIs, by default, use
the local multithreaded scheduler. Dask Bag, instead, relies on the local
multiprocessing scheduler.  All Dask data structure APIs, with the exception of
Dask Array and Dask DataFrame, were used in our experiments.

@@

 Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
 used Dask's default configuration that uses all the available memory on the
 instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
 Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave 5~GB
 for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
 default configuration for the master. We used the default configuration for worker
 memory management: at 60\% it spills data to disk, and 50\% of that amount is
 reserved for storage that is immune to eviction.

 One cloud instance did not host any worker and had a \SI{2}{\tera\byte} disk volume
 shared with the other instances using the Network File System (NFS) v4.
 This instance was also used for the Spark driver and master, for the
 Dask scheduler, and for job monitoring with the Spark and Dask user
 interfaces. For both Spark and Dask, spilled data is evicted to the NFS.

@@

We also used the dataset provided by the Consortium for Reliability and
Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
available on \href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and
functional images of 1,397 subjects acquired in 29 sites. We used all 3491 anatomical
images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on
average).

@@

%% Incrementation
\subsubsection{Incrementation}
We used an adaptation of the image incrementation pipeline used
in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
The application reads blocks of the BigBrain image from the shared file
system, increments the intensity value of each voxel by 1 to avoid caching
effects, sleeps for a configurable amount of time to emulate more
compute-intensive processing, repeats this process for a specified number of
iterations, and
finally writes the result as a NifTI image back to the shared file system.
This application allows us to study the behaviour of the engines when all
inputs are processed independently, in a map-only scenario (see
Figure~\ref{fig:tg-inc}). This mimics the behaviour of analyzing multiple
independent subjects in parallel.

@@

 Our second application calculates the histogram of the BigBrain image (see
 Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared file
 system, calculates intensity frequencies, aggregates the frequencies across the
 blocks, and finally writes the resulting histogram to the shared file system as a
 single \SI{766}{\kilo\byte} file. It follows the map-reduce
 paradigm, in which the final result is obtained from all the input blocks (see
 Fig.~\ref{fig:tg-histo}). This application requires data shuffling, incurring 
 additional performance costs compared to Algorithm~\ref{alg:incrementation}.
 The total amount of shuffled data is, however, limited to
 \SI{2.62}{\mega\byte} per blocks as it only consists of image histograms.
Two implementation are studied: pure Python and NumPy.

@@

For this experiment, we used the
\href{https://github.com/BIDS-Apps/example}{BIDS app example} map-reduce style 
application. The map phase, also called participant analysis, extracts the brain
tissues of a subject's 3D MRI using the FMRIB Software Library
(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting
image (\SI{2.47}{\mega\byte} on average) to the NFS. The reduce phase, or group
analysis, computes the volume of each brain and returns the average
volume, shuffling a total of \SI{8.6}{\giga\byte} image data.

While Incrementation and Histogram were implemented directly in Python, the 
BIDS App example requires an external library and is distributed as a Docker container image
(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
given its popularity for HPC and reproducibility, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
The image was preloaded on the NFS and singularity version 3.2.1\-1.el7 was installed
on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
export the path for FSL.

@@

The four parameters were varied in our experiments, as shown in
Table~\ref{tab:param}. We varied (1) the number of workers to assess the
scalability of the engines, (2) the number of BigBrain blocks in
Incrementation and Histogram to measure the effect of different IO patterns
and parallelization degrees, and (3) the number of iterations and sleep
delay in Incrementation to evaluate the effect of job length and number of
tasks.
It should be noted that increasing the number of blocks or iterations also
increases the total compute time of the application for a given sleep
delay. To avoid any potential external bias such as background load on the
network, we ran the experiments in a randomized order and cleared the page
cache of each worker before each execution.

For each run, we measured the application makespan as well as the cumulative 
data read, compute, data write, and engine overhead across all application
tasks. 
The overhead calculation for each CPU is the end time of its last processed task
minus the start time of the application. The summation of those results is the
total overhead.

@@

For all engines, the makespan is far from decreasing linearly with the
number of workers. The makespan
even increases between 4 and 8 workers. This is due to the high impact of
data transfers and engine overhead on the application.

@@

On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
slightly lower than Dask's, particularly when the number of workers increase.
Within Dask, Delayed and Futures have a higher overhead than Bag. However,
overhead differences are compensated by an increase in data transfer time,
as a reduced overhead increases the concurrency between data transfers.

@@

In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
For 30 blocks, Delayed and Futures have much higher overhead. This is because they are
thread-based and use multiple threads to compute a block which affect the calculation
of the overhead. Once again, the data transfer time reduces with more blocks however
the overhead time increases by a similar amount. This is not observed for 30 blocks
as the workers are not used at full capacity, i.e., some threads are idle. Finally,
the variability of the overhead increases with the number of blocks, which explains
the makespan variability mentioned previously.

@@

%% Sleep time
\subsection{Incrementation: Sleep delay}
Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
for different sleep delays. Overall, all engines again perform the same and scale
well with task duration. Spark is initially slower than the Dask APIs, however, it is
faster with an increased sleep delay. Also, within the Dask APIs, Dask Bag is
slower than the other two, but it is not considerable.

@@

%%% HISTOGRAM %%%
\subsection{Histogram: Number of workers}
Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
amounts of workers. Spark is significantly faster than Dask APIs. The difference
narrows as the number of workers increases but stays significant. Between Dask APIs
there is no substantial difference, however, Bag is slightly faster on average.

From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
shown. The compute time is significantly larger for Dask than for Spark. This is
potentially due to Python's GIL preventing Dask to parallelize the computation
over
multiple threads. Overall, the IO and overhead are comparable for all engines.

On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
when their number of workers varies, except for 8 workers where it increases slightly.
On the other hand, Spark total execution time keeps increasing; especially at 8
workers. This is because Dask engines benefit more from additional workers, in
this GIL-bounded scenario.

@@

\subsection{Histogram: Number of blocks}
Figure~\ref{fig:histo_ms_block} shows the makespan for different block sizes. 
Once more, Spark is significantly faster than Dask engines and Bag is slightly
faster than Delayed. Overall, the engines do not react to changes in block size.

In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
Note that at 30 blocks the workers' resources are not fully used since the
workers can process up to 64 tasks in parallel. Moreover, the overhead for
30 blocks is erroneous due to a lower amount of block than available threads.
This leads to thread based engines to have higher overhead. Given that, 
Spark has a much lower compute time which makes it significantly faster
than Dask engines. Overall, for all engines, lower block size results in lower
IO time but increases the overhead time. Again, this is due to large overhead
desynchronizing task thus reducing the NFS bottleneck.

@@

\subsection{Histogram: Gantt chart}
Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline Histogram
experiment. Spark overhead is distributed through all the execution while Bag
overhead happens mostly when a thread reads subsequent blocks and
Delayed overhead is concentrated after reading all the blocks.

% remove? - vhs
The reduce is done differently for the engines. Spark and Delayed compute
all the reduce tasks on one process (thread), however, Spark only computes it
on the driver process. On the other hand, Bag computes the reduce tasks on 
different workers and sends the information to the driver process afterward.

@@

%%% HISTOGRAM (NumPy) %%%
\subsection{Histogram (NumPy): Number of workers}
Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
Overall, the makespan for all engines does not change considerably with an increase
in the number of workers. This is likely a result of using a non-distributed
shared file system. There is no
substantial difference between all engines, although, Delayed is slightly faster.

@@

\subsection{Histogram (NumPy): Number of blocks}
Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
different block sizes. Overall, there is no substantial difference for
all engines and block sizes.

Figure~\ref{fig:histo_np_tt_block} shows the total time spent in each function.
Once more, since Delayed is thread-based its overhead at 30 blocks is seems much
higher that it really is. Considering the overhead error and that at 30 blocks
the workers only use half their core, the total time is substantially the same
for all block size. This is due to lower block sizes having a lower IO time,
however, it balances out by the resulting increase in overhead.

@@

\subsection{Histogram (NumPy): Gantt chart}
Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline for the
Histogram (NumPy) experiment. The overhead for Spark and Bag is mostly located
between read and compute tasks. On the other hand, Delayed
overhead is dispersed between all types of tasks. This causes Delayed to
overlap more compute and read tasks simultaneously amongst different workers.

@@

%%% BIDS EXAMPLE %%%
\subsection{BidsApp example: Number of workers}
Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when
varying the amount of workers. Overall, there is no substantial difference
between the engines. The makespan scales well with the increase in workers,
however, the scaling reduces considerably when reaching 8 workers. This is due
to an increase in overhead and an IO bottleneck caused parallel accesses to 
a non-distributed file system.

@@

\subsection{Effect of Network File System}
We used an NFS as it is a common approach to store a large amount of data. However,
it limits the available bandwidth since all workers interact with one disk. This
causes data-intensive applications to take a significant amount of time for IO. As we
expect, this effect is amplified when too many tasks simultaneously try to operate
with the NFS at the same time as
Fig.~\ref{fig:inc_gantt},\ref{fig:histo_gantt},\ref{fig:histo_np_gantt},\ref{fig:bids_gantt}
shows. A parallel file system, such as Lustre~\cite{lustre}, would likely improve
IO times of parallel applications simultaneously writing to shared storage.
